\textbf{Dynamic Hover Detection:\\Creating GUIs in Elm}

### 2013-12-15 paper/paper.tex
More revisions to 4

The first problem in implementing menus is twofold: how to position submenus on
the screen and how to hide submenus that should not be rendered. To get around
these problems, we replace submenus not shown on the screen with spacers. A
spacer is simply a blank rectangle with a specified width and height
(\texttt{spacer : Int -\textgreater{} Int -\textgreater{} Element}). For every
menu item, we create both the menu Element and a spacer. The spacer has the same
width as the Element, but a height of 1. We then switch between menu and the
spacer based on hover information. This helped in two ways. First of all, it
allowed us to choose between two options without creating a signal of signals.
Secondly, the spacer helped to align the submenu with its parent element.

@@

By replacing hidden elements with spacers, we are able to use the \texttt{flow}
primitive to position our elements instead of manually doing ``absolute''
positioning.  A submenu is rendered to the screen if one of three conditions
is met: the mouse is hovering upon its parent, the mouse is hovering upon it,
or the mouse is hovering upon any of its descendants.

@@

\texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
transition from True to False. By applying this function to every
hovering Boolean in the Menu structure, we give the mouse time to move
from a menu Element to its submenu before that submenu disappears.

A menu that does not detect mouse clicks is a fairly useless menu. There is no
primitive in Elm that can simply attach click detection to an element. The
closest primitive is \texttt{customButtons}. However, the implementation of
\texttt{customButtons} creates an extra dummy spacer which is rendered on top of
the element and therefore prevents us from changing the color of the menu
element when the mouse is hovering over it. To get around this issue, we used a
combination of \texttt{Mouse.clicks : Signal ()} (which fires an event whenever
the mouse clicks anywhere on the screen) and the hoverable information to
determine which element (if any) the mouse was hovering over.

\begin{verbatim}
clicksFromHover : Signal Bool -> Signal ()
clicksFromHover hover =
    lift (\_ -> ()) <| keepIf id False <| sampleOn Mouse.clicks hover
\end{verbatim}

@@

Interestingly enough, we were able to implement menus without explicitly
storing the state of the menu. There are two ways to remember state in
Elm code: \texttt{foldp}, or Elm's implementation of Arrowized FRP known
as \texttt{Automaton}. However, in our implementation we use the current
browser DOM state to remember our state. Instead of storing the menu
currently being displayed on the screen, we can derive it based on which
Element (if any) the mouse is currently hovering over. The elimination
of state from our code allows it to be much cleaner.

>>>>

Because \texttt{hoverablesJoin} was applied throughout the entire tree, it
contains every Element in every menu, and Booleans that indicate which ones to
display. Rather than omitting Elements associated with false, we replace these
hidden submenus with a spacer. A spacer is simply a blank rectangle with a
specified width and height (\texttt{spacer : Int -\textgreater{} Int
-\textgreater{} Element}). For our purposes, the spacer has the same width as
the Element, but a height of 1.\footnote{We were concerned a height of zero
would cause browser rendering issues.} We then switch between menu and the
spacer based on hover information. The spacer helped to align the submenu with
its parent element in a way a \texttt{Maybe Element} would not have done. By
replacing hidden elements with spacers, we are able to use the \texttt{flow}
primitive to position our elements instead of manually doing ``absolute''
positioning.

@@

A submenu is rendered to the screen if one of three conditions
is met: the mouse is hovering upon its parent, the mouse is hovering upon it,
or the mouse is hovering upon any of its descendants.

@@

Applying \texttt{extend} makes a \texttt{Signal Bool} wait a millisecond to
transition from true to false. Unlike a delay, true-to-false transitions still
occur immediately. By applying this function to every hovering Boolean in the
Menu structure, we give the mouse time to move from a menu Element to its
submenu before that submenu disappears. A millisecond was all that was needed.

A menu that does not detect mouse clicks is a fairly useless menu. There is no
primitive in Elm that can simply attach click detection to an element. To get
around this issue, we used a combination of \texttt{Mouse.clicks : Signal ()}
(which fires an event whenever the mouse clicks anywhere on the screen) and the
hoverable information to determine which element (if any) the mouse was hovering
over. (This implementation uses Elm's infix function application operator
\texttt{|>}, but it can be read as a pipeline.)

\begin{verbatim}
clicksFromHover : Signal Bool -> Signal ()
clicksFromHover hover =
    sampleOn Mouse.clicks hover
    |> keepIf id False
    |> lift (\_ -> ())
\end{verbatim}

@@

Interestingly, we were able to implement menus without explicitly storing the
state of the menu, i.e. without using \texttt{foldp} (or Elm's implementation of
Arrowized FRP known as \texttt{Automaton}). Rather we use the current browser
DOM state to remember state. Instead of storing the menu currently being
displayed on the screen, we can derive it based on which Element (if any) the
mouse is currently hovering over. The elimination of state from our code allows
it to be much cleaner.
